---
title: "Dirty Data Project"
author: "Geraldine A."
output:
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include=FALSE}

library(knitr)
library(readr)
library(here)
library(tidyverse)
library(kableExtra)
library(htmltools)

```

# Introduction

<div style="text-align: justify">
The dataset used in this analysis was collected from surveys conducted between 2015 and 2017, which gathered responses from individuals across various countries including the UK, USA, Spain, the Netherlands, and other European and Asian countries. The surveys asked participants to rate their level of satisfaction (on a scale of Joy, Despair, or Meh) when receiving different types of candies and treats during Halloween.

The purpose of this analysis is to identify patterns and insights within the dataset to better understand the preferences and satisfaction levels of individuals when it comes to Halloween candies and treats. By analyzing the data, we hope to gain a deeper understanding of the factors that influence people's satisfaction levels and how these factors differ across various countries and regions.

In the following sections, we will explore the dataset and present our findings in a clear and concise manner.

</div>

## Data 

<div style="text-align: justify">
For this analysis, I worked with three different data files that needed to be standardized, cleaned, and merged into a single file. The cleaning process involved removing any missing or incorrect data and handling inconsistencies in the formatting of the data.

One of the biggest challenges I encountered during the cleaning process was the lack of validation in the "country" column of the survey data. Respondents were allowed to enter any value they wanted in this column, leading to a large number of invalid or nonsensical responses. To address this issue, I implemented a process to identify and correct any erroneous data in the "country" column.

Despite these challenges, I was able to successfully prepare the data for analysis and extract valuable insights from it. The next sections will describe the analysis I performed on the data and the results I obtained.
</div>

## Analysis and visualisations



```{r echo=FALSE, include="FALSE"}
clean_data <- read_csv(here("data-clean/clean_data.csv"))

```

#### Total number of candy ratings given across the three years: 


```{r, echo=FALSE}
# Define column names to exclude from the count
cols_to_exclude <- c("how_old_are_you?", "trick_or_treat", "gender",
                    "country")

# Number of non-na ratings
num_of_non_rating <- clean_data %>%
  select(-one_of(cols_to_exclude)) %>%
  summarise_all(~ sum(!is.na(.))) %>%
  summarise(total_sum_ratings = sum(.))

kable(num_of_non_rating, format = "html", align = 'l') %>%
  kable_styling(full_width = T, position = "left")



```



#### Average age of people who are going out trick or treating?

```{r, echo=FALSE}

avg_age_that_who_going_out <- clean_data %>% 
  select(`how_old_are_you?`, trick_or_treat) %>% 
  filter(trick_or_treat == "yes") %>% 
  summarise(avg_age = round(mean(`how_old_are_you?`, na.rm = TRUE),0))

kable(avg_age_that_who_going_out, format = "html", align = 'l') %>%
  kable_styling(full_width = T) 

```



####  Average age of people who are not going trick or treating?

```{r, echo=FALSE}
avg_age_people_not_going_out <- clean_data %>% 
  select(`how_old_are_you?`, trick_or_treat) %>% 
  filter(trick_or_treat == "no") %>% 
  summarise(avg_age = round(mean(`how_old_are_you?`, na.rm = TRUE),0))

kable(avg_age_people_not_going_out, format = "html", align = 'l') %>%
  kable_styling(full_width = T, position = "left")

```



#### For each of joy, despair and meh, Which candy bar received the most of these ratings?

```{r, echo=FALSE}

# Create a candy_bar vector 

candy_bar_names <- c("butterfinger", "100_grand_bar", "any_full_sized_candy_bar", "snickers",
               "hershey's_dark_chocolate", "heath_bar", "hershey's_milk_chocolate",
               "kit_kat", "mars", "reggie_jackson_bar", "nestle_crunch", "milky_way",
               "tolberone_something_or_other","three_musketeers", "peanut_butter_bars",
               "twix", "mr._goodbar", "whatchamacallit_bars", "take_5")


candy_bar_rating <- clean_data %>%
  select(all_of(candy_bar_names)) %>%
  pivot_longer(cols = all_of(candy_bar_names),
         names_to = "candy_bar", values_to = "rating") %>%
  group_by(candy_bar, rating) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(candy_bar) %>%
  summarise(top_rating = if_else(all(is.na(count)), "People do not like this candy bar", 
                                         rating[which.max(count)]),
            rating_count = max(count, na.rm = TRUE)) %>%
  filter(!is.na(top_rating)) %>%
  arrange(desc(rating_count))



# set the height of the div
div_height <- "400px"

# add a scrollbar to the table
html_table <- paste0('<div style="height:', div_height, '; overflow-y: auto">', kable(candy_bar_rating[, -2], format = "html", align = 'l'), '</div>')

# display the table
HTML(html_table)

```


#### People rated Starburst as despair

```{r, echo=FALSE}

starburst_rate <- clean_data %>% 
  select(starburst) %>% 
  filter(starburst == "DESPAIR") %>% 
  count() %>% 
  rename(total_despair_count = n)

kable(starburst_rate, format = "html", align = 'l')


```

#### The most popular candy bar for each gender in the dataset

```{r, echo=FALSE}

# Define function to convert rating 
convert_rating <- function(rating) {
  case_when(
    rating == "DESPAIR" ~ -1,
    rating == "JOY" ~ 1,
    rating == "MEH" ~ 0
  )
}

candy_bar_data <- clean_data %>%
  select(all_of(candy_bar_names), gender)

# Convert ratings to numeric values
candy_data_numeric <- candy_bar_data %>%
  mutate_at(candy_bar_names, convert_rating)


most_popular_candy <- candy_data_numeric %>%
  group_by(gender) %>%
  summarise(across(everything(), ~sum(., na.rm = TRUE))) %>% 
  pivot_longer(cols = -gender, names_to = "candy_bar", values_to = "total_rating") %>% 
  filter(!is.na(gender) & total_rating != 0)


# set the height of the div
div_height <- "400px"

# add a scrollbar to the table
html_table <- paste0('<div style="height:', div_height, '; overflow-y: auto">', kable(most_popular_candy, format = "html", align = 'l'), '</div>')

# display the table
HTML(html_table)


```


#### The most popular candy bar in each year:
```{r, echo=FALSE}
candy_bar_data_by_year <- clean_data %>%
  select(all_of(candy_bar_names), year)

# Convert ratings to numeric values
candy_data_num <- candy_bar_data_by_year %>%
  mutate_at(candy_bar_names, convert_rating)


most_popular_candy_by_year <- candy_data_num %>%
  group_by(year) %>%
  summarise(across(everything(), ~ sum(., na.rm = TRUE))) %>%
  pivot_longer(cols = -year, names_to = "candy_bar", values_to = "total_rating") %>%
  filter(!is.na(year)) %>% 
  # filter(candy_bar != "any_full_sized_candy_bar") %>%
  arrange(year, desc(total_rating)) %>% 
  distinct(year, .keep_all = TRUE)


kable(most_popular_candy_by_year, format = "html", align = 'l')
```



#### The highest-rated candy bar among people in the US, Canada, UK, and other countries (excluding values with missing data):

```{r, echo=FALSE}
candy_bar_data_by_country <- clean_data %>%
  select(all_of(candy_bar_names), country)

# Convert ratings to numeric values
candy_data_country <- candy_bar_data_by_country %>%
  mutate_at(candy_bar_names, convert_rating)


most_popular_candy_by_country <- candy_data_country %>%
  group_by(country) %>%
  summarise(across(everything(), ~sum(., na.rm = TRUE))) %>%
  pivot_longer(cols = -country, names_to = "candy_bar", values_to = "total_rating") %>% 
  filter(!is.na(country)) %>% 
  # filter(candy_bar != "any_full_sized_candy_bar") %>%
  arrange(desc(total_rating)) %>% 
  distinct(country, .keep_all = TRUE) %>% 
  filter(!is.na(country) & total_rating != 0) 


# set the height of the div
div_height <- "400px"

# add a scrollbar to the table
html_table <- paste0('<div style="height:', div_height, '; overflow-y: auto">', kable(most_popular_candy_by_country, format = "html", align = 'l'), '</div>')

# display the table
HTML(html_table)

  
```




#### Top 3 most popular candies (per year / per gender)


```{r, echo=FALSE}

top_3_most_popular <- candy_data_num %>%
  group_by(year) %>%
  summarise(across(everything(), ~sum(., na.rm = TRUE))) %>%
  pivot_longer(cols = -year, names_to = "candy_bar", values_to = "total_rating") %>%
  filter(!is.na(year)) %>% 
  # filter(candy_bar != "any_full_sized_candy_bar") %>%
  arrange(year, desc(total_rating)) %>%
  group_by(year) %>%
  slice_max(order_by = total_rating, n = 3)



top_3_most_popular <- ggplot(top_3_most_popular, aes(x = candy_bar, total_rating, y = total_rating, fill = candy_bar)) +
  geom_col(position = "dodge", width = 0.7) +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.6, size = 10)) +
  theme(axis.text.y = element_text(size = 10)) +
  ggtitle("Top Three Candy Bars By Year") +
  xlab(NULL) +
  ylab("Total Rating") +
  theme(plot.title = element_text(hjust = 0.6)) +
  theme(legend.position = "right") +
  scale_color_brewer(palette = "Spectral") +
  scale_y_continuous(limits = c(0, max(top_3_most_popular$total_rating) * 1.1)) +
  facet_wrap(~year, scales = "free_x") +
  theme(plot.title = element_text(margin = margin(b = 20), face = "bold")) +
  theme(axis.title.x = element_text(margin = margin(t = 20))) +
  theme(axis.title.y = element_text(margin = margin(r = 20))) +
  theme(strip.text = element_text(size = 12, face = "bold")) +
  labs(fill = "Candy bar") +
  theme(axis.line.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank())
plot(top_3_most_popular)

```



### Ratings (per year / per gender)

```{r, echo=FALSE}

rating_by_gender_year <- clean_data %>%
  select(all_of(candy_bar_names), year, gender)

# Convert ratings to numeric values
rating_numeric <- rating_by_gender_year %>%
  mutate_at(candy_bar_names, convert_rating)


rating_by_gender <- rating_numeric %>%
  group_by(gender, year) %>%
  summarise(count = n(), .groups = "drop") %>%
  pivot_longer(cols = c(-gender, -year), names_to = "candy_bar", values_to = "total_rating") %>% 
  filter(!is.na(gender))

rating_by_gender <- ggplot(rating_by_gender, aes(x = gender, y = total_rating, fill = gender)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ year, nrow = 1) +
  labs(x = "Gender", y = "Total Rating", fill = "Candy Bar", 
       title = "Total Ratings by Gender for Each Year") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"), legend.position = "none") +
  theme(axis.text.y = element_text(size = 9)) +
  theme(axis.text.x = element_text(size = 9)) +
  theme(strip.text = element_text(size = 12, face = "bold"))
plot(rating_by_gender)

```



#### Missing values (per year / per gender / per country)

```{r, echo=FALSE}

# Missing data by year


missing_data_by_year <- clean_data %>%
  group_by(year) %>% 
  summarise(across(everything(), ~sum(is.na(.)))) %>% 
  mutate(total = rowSums(select(., -year))) %>% 
  select(year, total) %>% 
  ggplot(aes(x = year, y = total, fill = factor(year))) +
  geom_bar(stat = "identity") +
  labs(title = "Missing Data by Year", x = "Year", y = "Total Missing Data", fill = "Year") +
  scale_fill_brewer(palette = "Dark2")+
  scale_y_continuous(labels = scales::comma_format(scale = 1)) 

missing_data_by_year
```



#### Missing data by Gender
```{r, echo=FALSE}
missing_data_by_gender <- clean_data %>%
  group_by(gender) %>% 
  summarise(across(everything(), ~sum(is.na(.)))) %>% 
  mutate(total = rowSums(select(., -gender))) %>% 
  select(gender, total) %>% 
  filter(!is.na(gender)) %>% 
  ggplot(aes(x = gender, y = total, fill = factor(gender))) +
  geom_bar(stat = "identity") +
  labs(title = "Missing Data by Gender", x = "Year", y = "Total Missing Data", fill = "Gender") +
  scale_fill_brewer(palette = "Set1")+
  scale_y_continuous(labels = scales::comma_format(scale = 1)) 

plot(missing_data_by_gender)

```



### Missing data by country:
```{r, echo=FALSE}
missing_data_by_country <- clean_data %>%
  group_by(country) %>% 
  summarise(across(everything(), ~sum(is.na(.)))) %>% 
  mutate(total = rowSums(select(., -country))) %>% 
  select(country, total) %>% 
  filter(!is.na(country)) 

missing_data_by_country %>%
  mutate(country = fct_reorder(country, total)) %>%
  ggplot(aes(x = total, y = country)) +
  geom_col(fill = "steelblue") +
  labs(title = "Missing Data by Country", x = "Total Missing Data", y = "Country") +
  scale_fill_brewer(palette = "Paired") +
  scale_x_continuous(labels = scales::comma_format(scale = 1)) +
  coord_cartesian(xlim = c(0, max(missing_data_by_country$total, na.rm = TRUE) * 1.1))


```

# Conclusion
<div style="text-align: justify">
After analyzing the candy surveys dataset, we can conclude that the average age of people who go out for trick or treating is 35, while the average age of people who do not go out is 39.

The dataset also showed that people rated the candy bars based on three categories - joy, despair, and meh. Based on this rating system, the most enjoyed candy bar was any full-sized candy bar (7246 ratings), followed by Kit Kat (6912 ratings), and Twix (6684 ratings).

Furthermore, we discovered that the most popular candy bars for each gender were slightly different. Females preferred any full-sized candy bar (855 ratings), Butterfinger, and 100 Grand Bar, while males preferred the same candy bars but with a higher rating for any full-sized candy bar (1533 ratings).

In terms of yearly popularity, any full-sized candy bar was the most popular candy bar for 2015, 2016, and 2017, with 4392, 982, and 1497 ratings respectively.

Finally, we found that the highest-rated candy bar by country was any full-sized candy bar, which was the top-rated candy bar for the USA, Canada, the UK, and Germany.

It is worth noting that there were many missing data points in the dataset, with the majority of them coming from the USA (101834 NAs), followed by Canada (13179 NAs) and the UK (3100 NAs). Overall, this analysis provides valuable insights into candy preferences and trick-or-treating habits among different demographics.
<div>



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
