---
title: "Analysis"
output: html_notebook
---

```{r}
library(dplyr)
library(here)
library(styler)
library(tidyr)
library(readxl)
library(ggplot2)


```

```{r}

here::here()

clean_data <- read_csv(here("data-clean/clean_data.csv"))

```



# Analysis questions

## 1. What is the total number of candy ratings given across the three years. 
## (Number of candy ratings, not the number of raters. Don’t count missing values)


```{r}

# Define column names to exclude from the count
cols_to_exclude <- c("how_old_are_you?", "trick_or_treat", "gender",
                    "country")

# Number of non-na ratings
num_of_non_rating <- clean_data %>%
  select(-one_of(cols_to_exclude)) %>%
  summarise_all(~ sum(!is.na(.))) %>%
  summarise(total_sum_ratings = sum(.))

```
## 2. What was the average age of people who are going out trick or treating?

```{r}
avg_age_that_who_going_out <- clean_data %>% 
  select(`how_old_are_you?`, trick_or_treat) %>% 
  filter(trick_or_treat == "yes") %>% 
  summarise(avg_age = round(mean(`how_old_are_you?`, na.rm = TRUE),0))

```
## 3. What was the average age of people who are not going trick or treating?

```{r}
avg_age_people_not_going_out <- clean_data %>% 
  select(`how_old_are_you?`, trick_or_treat) %>% 
  filter(trick_or_treat == "no") %>% 
  summarise(avg_age = round(mean(`how_old_are_you?`, na.rm = TRUE),0))
```
## 4. For each of joy, despair and meh, which candy bar received the most of these ratings?


```{r}
# Create a candy_bar vector 

candy_bar_names <- c("butterfinger", "100_grand_bar", "any_full_sized_candy_bar", "snickers",
               "hershey's_dark_chocolate", "heath_bar", "hershey's_milk_chocolate",
               "kit_kat", "mars", "reggie_jackson_bar", "nestle_crunch", "milky_way",
               "tolberone_something_or_other","three_musketeers", "peanut_butter_bars",
               "twix", "mr._goodbar", "whatchamacallit_bars", "take_5")


candy_bar_rating <- clean_data %>%
  select(all_of(candy_bar_names)) %>%
  pivot_longer(cols = candy_bar_names,
         names_to = "candy_bar", values_to = "rating") %>%
  group_by(candy_bar, rating) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(candy_bar) %>%
  summarise(total_count = sum(count, na.rm = TRUE),
            most_common_rating = if_else(all(is.na(count)), "People do not like this candy bar", 
                                         rating[which.max(count)]),
            most_common_rating_count = max(count, na.rm = TRUE)) %>%
  filter(!is.na(most_common_rating))

candy_bar_rating %>% 
  arrange(desc(most_common_rating_count))

```

## 5. How many people rated Starburst as despair?

```{r}

starburst_rate <- clean_data %>% 
  select(starburst) %>% 
  filter(starburst == "DESPAIR") %>% 
  count() %>% 
  rename(total_despair_count = n)

```
## 6. What was the most popular candy bar by this rating system for each gender in the dataset?
## Count despair as -1, joy as +1, and meh as 0.

```{r}

# Define function to convert rating 
convert_rating <- function(rating) {
  case_when(
    rating == "DESPAIR" ~ -1,
    rating == "JOY" ~ 1,
    rating == "MEH" ~ 0
  )
}

candy_bar_data <- clean_data %>%
  select(candy_bar_names, gender)

# Convert ratings to numeric values
candy_data_numeric <- candy_bar_data %>%
  mutate_at(candy_bar_names, convert_rating)


most_popular_candy <- candy_data_numeric %>%
  group_by(gender) %>%
  summarise(across(everything(), sum, na.rm = TRUE)) %>% 
  pivot_longer(cols = -gender, names_to = "candy_bar", values_to = "total_rating") %>% 
  filter(!is.na(gender) & total_rating != 0)

most_popular_candy


```

## 7. What was the most popular candy bar in each year?
```{r}
candy_bar_data_by_year <- clean_data %>%
  select(candy_bar_names, year)

# Convert ratings to numeric values
candy_data_num <- candy_bar_data_by_year %>%
  mutate_at(candy_bar_names, convert_rating)


most_popular_candy_by_year <- candy_data_num %>%
  group_by(year) %>%
  summarise(across(everything(), sum, na.rm = TRUE)) %>%
  pivot_longer(cols = -year, names_to = "candy_bar", values_to = "total_rating") %>%
  filter(!is.na(year)) %>% 
  filter(candy_bar != "any_full_sized_candy_bar") %>%
  arrange(year, desc(total_rating)) %>% 
  distinct(year, .keep_all = TRUE)

most_popular_candy_by_year
```

## 8. What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?
```{r}

candy_bar_data_by_country <- clean_data %>%
  select(candy_bar_names, country)
# Convert ratings to numeric values
candy_data_country <- candy_bar_data_by_country %>%
  mutate_at(candy_bar_names, convert_rating)


most_popular_candy_by_country <- candy_data_country %>%
  group_by(country) %>%
  summarise(across(everything(), sum, na.rm = TRUE)) %>%
  pivot_longer(cols = -country, names_to = "candy_bar", values_to = "total_rating") %>% 
  filter(!is.na(country)) %>% 
  filter(candy_bar != "any_full_sized_candy_bar") %>%
  arrange(desc(total_rating)) %>% 
  distinct(country, .keep_all = TRUE) %>% 
  filter(!is.na(country) & total_rating != 0) 
  
```

## Visualizations

## Top 3 most popular candies (per year / per gender / per country)


```{r}

top_three_candy_bar <- candy_data_num %>%
  group_by(year) %>%
  summarise(across(everything(), sum, na.rm = TRUE)) %>%
  pivot_longer(cols = -year, names_to = "candy_bar", values_to = "total_rating") %>%
  filter(!is.na(year)) %>% 
  filter(candy_bar != "any_full_sized_candy_bar") %>%
  arrange(year, desc(total_rating)) %>%
  group_by(year) %>%
  slice_max(order_by = total_rating, n = 3)



ggplot(top_three_candy_bar, aes(x = candy_bar, total_rating, y = total_rating, fill = candy_bar)) +
  geom_col(position = "dodge", width = 0.5) +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5, size = 11)) +
  theme(axis.text.y = element_text(size = 11)) +
  ggtitle("Top Three Candy Bars By Year") +
  xlab("Candy Bar") +
  ylab("Total Rating") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(legend.position = "right") +
  scale_color_brewer(palette = "Spectral") +
  scale_y_continuous(limits = c(0, max(top_three_candy_bar$total_rating) * 1.1)) +
  facet_wrap(~year, scales = "free_x") +
  theme(plot.title = element_text(margin = margin(b = 20), face = "bold")) +
  theme(axis.title.x = element_text(margin = margin(t = 20))) +
  theme(axis.title.y = element_text(margin = margin(r = 20))) +
  theme(strip.text = element_text(size = 12, face = "bold")) + 
  labs(fill = "Candy bar")




```


## Proportion of ratings (per year / per gender)

```{r}

rating_by_gender_year <- clean_data %>%
  select(candy_bar_names, year, gender)

# Convert ratings to numeric values
rating_numeric <- rating_by_gender_year %>%
  mutate_at(candy_bar_names, convert_rating)


rating_by_gender <- rating_numeric %>%
  group_by(gender, year) %>%
  summarise(count = n(), .groups = "drop") %>%
  pivot_longer(cols = c(-gender, -year), names_to = "candy_bar", values_to = "total_rating") %>% 
  filter(!is.na(gender))

ggplot(rating_by_gender, aes(x = gender, y = total_rating, fill = gender)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ year, nrow = 1) +
  labs(x = "Gender", y = "Total Rating", fill = "Candy Bar", 
       title = "Total Ratings by Gender for Each Year") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"), legend.position = "none") +
  theme(axis.text.y = element_text(size = 11)) +
  theme(axis.text.x = element_text(size = 11)) +
  theme(strip.text = element_text(size = 12, face = "bold"))


```
## Missing values (per year / per gender / per country)

```{r}

# Missing data by year


missing_data_by_year <- clean_data %>%
  group_by(year) %>% 
  summarise(across(everything(), ~sum(is.na(.)))) %>% 
  mutate(total = rowSums(select(., -year))) %>% 
  select(year, total) %>% 
  ggplot(aes(x = year, y = total, fill = factor(year))) +
  geom_bar(stat = "identity") +
  labs(title = "Missing Data by Year", x = "Year", y = "Total Missing Data", fill = "Year") +
  scale_fill_brewer(palette = "Dark2")+
  scale_y_continuous(labels = scales::comma_format(scale = 1)) 

missing_data_by_year



```

```{r}
# Missing data by Gender

missing_data_by_gender <- clean_data %>%
  group_by(gender) %>% 
  summarise(across(everything(), ~sum(is.na(.)))) %>% 
  mutate(total = rowSums(select(., -gender))) %>% 
  select(gender, total) %>% 
  filter(!is.na(gender)) %>% 
  ggplot(aes(x = gender, y = total, fill = factor(gender))) +
  geom_bar(stat = "identity") +
  labs(title = "Missing Data by Gender", x = "Year", y = "Total Missing Data", fill = "Gender") +
  scale_fill_brewer(palette = "Set1")+
  scale_y_continuous(labels = scales::comma_format(scale = 1)) 

missing_data_by_gender

```
```{r}
# Missing data by country 

missing_data_by_country <- clean_data %>%
  group_by(country) %>% 
  summarise(across(everything(), ~sum(is.na(.)))) %>% 
  mutate(total = rowSums(select(., -country))) %>% 
  select(country, total) %>% 
  filter(!is.na(country)) 



missing_data_by_country %>%
  mutate(country = fct_reorder(country, total)) %>%
  ggplot(aes(x = total, y = country)) +
  geom_col(fill = "steelblue") +
  labs(title = "Missing Data by Country", x = "Total Missing Data", y = "Country") +
  scale_fill_brewer(palette = "Paired") +
  scale_x_continuous(labels = scales::comma_format(scale = 1))


```
 
